Efficiently stable presentations from error-correcting codes

### 2023-07-17 efficient_stability.tex
added parity check formulation of code concatenation

\hnote{added the parity check version of this test:} A parity check matrix $h_{\RM} \in \F^{m q^{m+1} \times q^m}$ for $\code_{\RM}$ is as follows. Identify the rows with triples $(u,j,t) \in \F_q^m \times \{1,\ldots,m\} \times (\F_q \setminus \{t_0,\ldots,t_d\})$. The $(u,j,t)$-th row of $h_{\RM}$ is the vector in $\F_q^m$ that has $\alpha_{u,u+te_j,i}$ in the coordinate indexed by $u + t_i e_j$, and $-1$ in the coordinate indexed by $u + te_j$. Rubinfeld and Sudan \hnote{need to double check reference} showed that $h_{\RM}$ is indeed a parity check matrix for $\code_{\RM}$, and... \hnote{complete later}

@@

To sample from $\nu'_1$, sample $S\sim\nu$ and $\gamma\in \F_2^t$ uniformly at random. If $M$ performed the check $v_S \cdot u_S=0$, then perform the check 
\begin{align*}
0 &= \tr\big(( v_S \cdot u_S) \gamma\big)\\
&= \sum_{j\in S} \tr( v_{j} u_j \gamma) \\
&= \sum_{j\in S} \kappa(\gamma v_{j}) \cdot \kappa(u_j)\;.
\end{align*}
This check can be performed by evaluating $u$ at the coordinates $(j,\kappa(\gamma  v_{j}))$ for $j\in S$ and summing the results in $\F_2$.

@@

\hnote{I'm a little confused by the description of this local tester... This seems to just make $r$ queries, right (but it says $rq$ above)? And it feels like to ``simulate'' the check from $\code$, one has to make $2q$ random queries to the Hadamard code to decode the symbol from $\F_q$, right?}\tnote{You're right, it's just $r$ queries. I don't think that we need to recover the entire $\F_q$ symbols, because we're checking a linear equation. What we're effectively doing (by multiplying by $\gamma$ and taking the trace) is checking a ``random projection'' of the equation. Like a random linear combination of the $t$ equations implied by one equation over $\F_q$. We definitely don't want any $q$-dependence in the tester}

>>>>

\hnote{added the parity check version of this test:} A parity check matrix $h_{\RM} \in \F^{S \times q^m}$ for $\code_{\RM}$ where $S = q^m \times m \times (q - d-1)$ is as follows. Identify the rows with triples $(u,j,t) \in \F_q^m \times \{1,\ldots,m\} \times (\F_q \setminus \{t_0,\ldots,t_d\})$. The $(u,j,t)$-th row of $h_{\RM}$ is the vector in $\F_q^m$ that has $\alpha_{u,u+te_j,i}$ in the coordinate indexed by $u + t_i e_j$, and $-1$ in the coordinate indexed by $u + te_j$. Babai, Fortnow, and Lund essentially showed that \hnote{need to double check reference} $h_{\RM}$ is indeed a parity check matrix for $\code_{\RM}$. \hnote{complete later}

@@

To sample from $\nu'_1$, sample $S\sim\nu$ and $\gamma\in \F_2^t$ uniformly at random \hnote{type-checking, $\gamma$ should be in $\F_q$, I think}. If $M$ performed the check $v_S \cdot u_S=0$, then perform the check 
\begin{align*}
0 &= \tr\big(( v_S \cdot u_S) \gamma\big)\\
&= \sum_{j\in S} \tr( v_{j} u_j \gamma) \\
&= \sum_{j\in S} \kappa(\gamma v_{j}) \cdot \kappa(u_j)\;.
\end{align*}
This check can be performed by evaluating $u$ at the coordinates $(j,\kappa(\gamma  v_{j}))$ for $j\in S$ and summing the results in $\F_2$.

@@

\hnote{added the parity check version of the tester: } Let $M = (h,\nu)$ denote an $r$-local $\delta$-tester for $\code$, where $h \in \F_q^{\ell \times n}$ is a parity check matrix for $\code$. We now describe an $r$-local $(\cdots)$-tester $M' = (h',\nu')$ for $\code'$. Define $h'$ to be the block matrix $\begin{bmatrix} A \\ B \end{bmatrix}$ where 
\begin{itemize}
	\item $A \in \F^{nT^2 \times nT}$ is itself a block-diagonal matrix where the diagonal blocks are the $T^2 \times T$ parity check matrix for the Hadamard code. 
	In other words, $A$ can be viewed as $I_{n\times n} \otimes h_\had$ where $h_\had \in \F_2^{T^2 \times T}$ is the parity check matrix for the Hadamard code. 
	\item $B \in \F^{\ell T \times nT}$ is viewed as having the rows indexed by pairs $(p,\gamma) \in \{1,\ldots,\ell\} \times \F_q$ and columns indexed by pairs $(i,x) \in \{1,\ldots,n\} \times \F_2^t$. The entry in row $(p,\gamma)$ and column $(i,x)$ is $1$ if and only if $h_{pi} \neq 0$ and $x = \kappa(\gamma h_{pi})$. \hnote{double check this}
\end{itemize}
It is easy to check that $h'$ is a parity check matrix for $\code'$. Finally, define the distribution $\nu'$ to be... \hnote{to be continued}
